---
editor_options: 
  chunk_output_type: console
---

# Light availability

In this script, we examine differences in vocal activity between dawn and dusk as a function of light availability. The expectation is that species would call much earlier in the day, closer to sunrise compared to later in the day when one could spend time foraging.  

## Install necessary libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(ggstatsplot)
library(suncalc)
library(lutz)
library(stringr)
library(ggrepel)
library(ggpubr)
```

## Load acoustic data and species scientific names data
```{r}
acoustic_data <- read.csv("results/acoustic_data.csv")
sites <- read.csv("data/list-of-sites.csv")
species_codes <- read.csv("data/species-annotation-codes.csv")
```

## Extract light availability for each date

```{r}
# add longitude and latitude to acoustic_data
acoustic_data <- left_join(acoustic_data, sites[,c(2,4,5)],
                   by = "site_id") 
acoustic_data$date <- lubridate::ymd(acoustic_data$date)
names(acoustic_data)[c(10,11)] <- c("lon","lat")

# find out what time zone needs to be provided for the sunlight calculations
acoustic_data$tz <- tz_lookup_coords(lat = acoustic_data$lat,
                                     lon = acoustic_data$lon,
                                     method = "accurate",
                                     warn = FALSE)

# extract nauticalDawn, nauticalDusk, sunrise and sunset times
light_data <- getSunlightTimes(data = acoustic_data,
                          keep = c("sunrise","sunset",
                                   "nauticalDawn", "nauticalDusk"),
                          tz = "Asia/Kolkata") %>%
  distinct(.)

# strip dates from new columms and keep only time
light_data$sunrise <- as_hms(light_data$sunrise)
light_data$sunset <- as_hms(light_data$sunset)
light_data$nauticalDawn <- as_hms(light_data$nauticalDawn)
light_data$nauticalDusk <- as_hms(light_data$nauticalDusk)

# format the start_time column in the acoustic data to keep it as the same format as light_data
acoustic_data <- acoustic_data %>%
   mutate(across(start_time, str_pad, width = 6, pad = "0"))
acoustic_data$start_time<- format(strptime(acoustic_data$start_time, 
                         format = "%H%M%S"), format = "%H:%M:%S")
acoustic_data$start_time <- as_hms(acoustic_data$start_time)

# summarize detections of species for every 15-min window
acoustic_data <- acoustic_data %>%
  group_by(site_id, date, start_time, time_of_day, eBird_codes,
           lon, lat) %>%
  summarise(detections = sum(number)) %>%
  ungroup()

# join the two datasets
acoustic_data <- left_join(acoustic_data, light_data,
                   by = c("date","lon","lat"))

# subtract times from sunrise, sunset, nauticalDawn and nauticalDusk from start_time of acoustic detections
acoustic_data <- acoustic_data %>%
  mutate(time_from_dawn = as.numeric((start_time - nauticalDawn), 
                                     units = "hours")) %>%
  mutate(time_from_sunrise = as.numeric((start_time - sunrise), 
                                     units = "hours")) %>%
  mutate(time_to_dusk = as.numeric((nauticalDusk-start_time), 
                                     units = "hours")) %>%
  mutate(time_to_sunset = as.numeric((sunset-start_time), 
                                     units = "hours"))
```

## Model acoustic detections as a function of light availability  

Here, we choose times of day as proxies for light availability (ie. nautical dawn, dusk for example).  

```{r}
trial <- acoustic_data %>%
  filter(eBird_codes == "crbsun2") %>%
  filter(time_of_day == "dusk") %>%
  ggplot(., aes(x=time_to_dusk,
                y= detections)) +
  geom_point(shape = 21, colour = "black", 
             fill = "white", size = 2, stroke = 1)+ 
  geom_smooth(method="lm", se=TRUE, fullrange=FALSE, level=0.95,linetype="solid") +  
  theme_bw() +
  stat_regline_equation(label.x = 1, aes(label = ..rr.label..),
                    label.y = 20,
                      size = 8) +
  labs(y="\n nDetections", 
       x="Time from Dawn\n") +
  #geom_text_repel(aes(label = common_name), family = "Century Gothic",
  #                fontface = "italic") +
  theme(text = element_text(family = "Century Gothic", size = 18, face = "bold"),plot.title = element_text(family = "Century Gothic",
      size = 18, face = "bold"),
      plot.subtitle = element_text(family = "Century Gothic", 
      size = 15, face = "bold",color="#1b2838"),
      axis.title = element_text(family = "Century Gothic",
      size = 18, face = "bold"))

trial

```


