---
editor_options: 
  chunk_output_type: console
---
 
  
# Phylogenetic Generalized Least Squares (PGLS) regressions
  
In this script, we examine which predictor/hypothesis (light availability, peak frequency, trophic niche, territoriality, and sociality) best explains the variation in vocal activity between dawn and dusk. We use linear regresssions while controlling for potential phylogenetic signals/relationships.

## Loading necessary libraries
```{r}
library(ape)
library(nlme)
library(geiger)
library(phytools)
library(phangorn)
library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(stringr)
library(vegan)
library(scico)
library(data.table)
library(extrafont)
library(suncalc)
library(lutz)
library(ggpmisc)
library(ggpubr)
library(hms)
library(phyr)
library(sjPlot)
```

## Loading necessary data
```{r}
acoustic_data <- read.csv("results/acoustic_data.csv")
species_codes <- read.csv("data/species-annotation-codes.csv")
trait <- read.csv("data/species-trait-dat.csv")
territoriality <- read.csv("data/territoriality-data.csv")
sociality <- read.csv("data/sociality-data.csv")
freq <- read.csv("data/frequency-data.csv")
sites <- read.csv("data/list-of-sites.csv")
vegetation <- read.csv("data/vegetation-data.csv")
tree <- read.nexus("data/birdtree.nex") #obtained 100 trees using Hackett all species 
```

## Getting the maximum credibility tree from the nexus file
```{r}
tree <- mcc(tree)
```

## Vocal Activity

We use normalized percent detections (percent detections which account for sampling effort) as a response variable in our PGLS analysis.
```{r}
# sampling effort by time_of_day
effort <- acoustic_data %>%
    dplyr::select(site_id, date, time_of_day) %>%
    distinct() %>%
    arrange(time_of_day) %>%
    count(time_of_day) %>%
    rename(., nVisits = n)

# Above, we note that we had sampled ~293 site-date combinations at dawn, while ~245 site-date combinations were sampled at dusk

# total number of acoustic detections summarized across every 10-s audio file
# here, we estimate % detections at dawn and dusk, while accounting for sampling effort
vocal_act <- acoustic_data %>%
  group_by(time_of_day, eBird_codes) %>%
  summarise(detections = sum(number)) %>%
  left_join(., species_codes[,c(1,2,5)], 
                       by = "eBird_codes") %>%
   group_by(eBird_codes) %>%
  mutate(total_detections =sum(detections)) %>%
  mutate(percent_detections = (detections/total_detections)*100) %>%
  ungroup() 

## accouting for sampling effort and normalizing data
vocal_act <- vocal_act %>%
  left_join(., effort, by = "time_of_day") %>%
  mutate(normalized_detections = detections/nVisits) %>%
  group_by(eBird_codes) %>%
  mutate(total_normalized_detections = sum(normalized_detections)) %>%
  mutate(percent_normalized_detections = (normalized_detections/total_normalized_detections)*100) %>%
  ungroup()
```

## Light availability  

Here, we prepare a predictor for light availability or time to darkness for the PGLS analysis
```{r}
# add longitude and latitude to acoustic_data
acoustic_data <- left_join(acoustic_data, sites[,c(2,4,5)],
                   by = "site_id") 
acoustic_data$date <- lubridate::ymd(acoustic_data$date)
names(acoustic_data)[c(10,11)] <- c("lon","lat")

# find out what time zone needs to be provided for the sunlight calculations
acoustic_data$tz <- tz_lookup_coords(lat = acoustic_data$lat,
                                     lon = acoustic_data$lon,
                                     method = "accurate",
                                     warn = FALSE)

# extract nauticalDawn, nauticalDusk, sunrise and sunset times
light_data <- getSunlightTimes(data = acoustic_data,
                          keep = c("sunrise","sunset",
                                   "nauticalDawn", "nauticalDusk"),
                          tz = "Asia/Kolkata") %>% distinct(.)

# strip dates from new columms and keep only time
light_data$sunrise <- as_hms(light_data$sunrise)
light_data$sunset <- as_hms(light_data$sunset)
light_data$nauticalDawn <- as_hms(light_data$nauticalDawn)
light_data$nauticalDusk <- as_hms(light_data$nauticalDusk)

# format the start_time column in the acoustic data to keep it as the same format as light_data
acoustic_data <- acoustic_data %>%
   mutate(across(start_time, str_pad, width = 6, pad = "0"))
acoustic_data$start_time<- format(strptime(acoustic_data$start_time, 
                         format = "%H%M%S"), format = "%H:%M:%S")
acoustic_data$start_time <- as_hms(acoustic_data$start_time)

# summarize detections of species for every 15-min window
acoustic_data <- acoustic_data %>%
  group_by(site_id, date, start_time, time_of_day, eBird_codes,
           lon, lat) %>%
  summarise(detections = sum(number)) %>%
  ungroup()

# join the two datasets
acoustic_data <- left_join(acoustic_data, light_data,
                   by = c("date","lon","lat"))

# subtract times from sunrise, sunset, nauticalDawn and nauticalDusk from start_time of acoustic detections
acoustic_data <- acoustic_data %>%
  mutate(time_from_dawn = as.numeric((start_time - nauticalDawn), 
                                     units = "hours")) %>%
  mutate(time_from_sunrise = as.numeric((start_time - sunrise), 
                                     units = "hours")) %>%
  mutate(time_to_dusk = as.numeric((nauticalDusk-start_time), 
                                     units = "hours")) %>%
  mutate(time_to_sunset = as.numeric((sunset-start_time), 
                                     units = "hours"))

# add species scientific name to this data
acoustic_data <- acoustic_data %>%
  left_join(., species_codes[,c(1, 5)], 
                       by = "eBird_codes")

# binding dawn and dusk times together to get at median time/light availability at dawn 
dawn <- acoustic_data %>%
  group_by (eBird_codes) %>%
  filter(time_of_day == "dawn") %>%
  dplyr::select(time_from_dawn, time_of_day) %>%
  rename(., time_from_startTime = time_from_dawn) %>%
  summarise(median_startTime =median(time_from_startTime)) %>%
  mutate (time_of_day = 'dawn')

dusk <- acoustic_data %>%
  group_by (eBird_codes) %>%
  filter(time_of_day == "dusk") %>%
  dplyr::select(time_to_dusk, time_of_day) %>%
  rename(., time_from_startTime = time_to_dusk) %>%
 summarise(median_startTime =median(time_from_startTime)) %>%
  mutate (time_of_day = 'dusk')

light <- bind_rows (dawn, dusk)# this dataframe gives us the median time from dawn and time to dusk for each species

vocal_act <- vocal_act %>%
  left_join(light, by = c('eBird_codes', 'time_of_day')) #adding median time from dawn and time to dusk to vocal activity data
```

## Territoriality  

We prepare territoriality as a predictor for the PGLS analysis
```{r}
territoriality <- territoriality %>% dplyr::select (Species, Territory)
colnames (territoriality) <- c("scientific_name", "territory")

# Updating the scientific names of following species as per our dataset:
# Flame-throated bulbul- Rubigula gularis (earlier Pycnonotus gularis)
# Black eagle- Ictinaetus malaiensis (earlier Ictinaetus malayensis)
# Brown-capped pygmy woodpecker- Yungipicus nanus (earlier Dendrocopos nanus)
# Malabar barbet- Psilopogon malabaricus (earlier Megalaima malabarica)
# Dark-fronted babbler- Dumetia atriceps (earlier Rhopocichla atriceps)
# Greater flameback- Chrysocolaptes guttacristatus (earlier Chrysocolaptes lucidus)
# Indian blue robin- Larvivora brunnea (earlier Luscinia brunnea)
# Indian yellow tit- Machlolophus aplonotus (earlier Parus aplonotus)
# Jungle babbler- Argya striata (earlier Turdoides striata)
# Orange-headed thrush- Geokichla citrina (earlier Zoothera citrina)
# Rufous babbler- Argya subrufa (earlier Turdoides subrufa)
# Rufous woodpecker- Micropternus brachyurus (earlier Celeus brachyurus)
# Rusty-tailed flycatcher- Ficedula ruficauda (earlier Muscicapa ruficauda
# Spot-bellied eagle owl- Ketupa nipalensis (earlier Bubo nipalensis)
# Spotted dove- Spilopelia chinensis (earlier Stigmatopelia chinensis)
# Thick-billed warbler- Arundinax aedon (earlier Acrocephalus aedon)
# White-bellied flycatcher- Cyornis pallidipes (earlier Cyornis pallipes)
# White-cheeked barbet- Psilopogon viridis (earlier Megalaima viridis)
# Wayanad laughingthrush- Pterorhinus delesserti (earlier Garrulax delesserti)
# Yellow-browed bulbul- Iole indica (earlier Acritillas indica)

territoriality <- territoriality %>% mutate(scientific_name = recode (scientific_name, 'Pycnonotus gularis' = 'Rubigula gularis', 'Ictinaetus malayensis' = 'Ictinaetus malaiensis', 'Dendrocopos nanus' = 'Yungipicus nanus', 'Megalaima malabarica' = 'Psilopogon malabaricus', 'Rhopocichla atriceps' = 'Dumetia atriceps', 'Chrysocolaptes lucidus' = 'Chrysocolaptes guttacristatus', 'Luscinia brunnea' = 'Larvivora brunnea', 'Parus aplonotus' = 'Machlolophus aplonotus', 'Turdoides striata' = 'Argya striata', 'Zoothera citrina' = 'Geokichla citrina', 'Turdoides subrufa' = 'Argya subrufa', 'Celeus brachyurus' = 'Micropternus brachyurus', 'Muscicapa ruficauda' = 'Ficedula ruficauda', 'Bubo nipalensis' = 'Ketupa nipalensis', 'Stigmatopelia chinensis' = 'Spilopelia chinensis', 'Acrocephalus aedon' = 'Arundinax aedon', 'Cyornis pallipes' = 'Cyornis pallidipes', 'Megalaima viridis' = 'Psilopogon viridis', 'Garrulax delesserti' = 'Pterorhinus delesserti', 'Acritillas indica' = 'Iole indica')) 

vocal_act <- vocal_act %>%
  left_join(territoriality, by = 'scientific_name') 
```

## Sociality  

We prepare sociality as a predictor for the PGLS analysis  
```{r}
sociality <- sociality %>% dplyr::select (Species, Communal)
colnames (sociality) <- c("scientific_name", "sociality")

# Updating the scientific names of following species as per our dataset: see list in the code chunk for Territoriality

sociality <- sociality %>% mutate(scientific_name = recode (scientific_name, 'Pycnonotus gularis' = 'Rubigula gularis', 'Ictinaetus malayensis' = 'Ictinaetus malaiensis', 'Dendrocopos nanus' = 'Yungipicus nanus', 'Megalaima malabarica' = 'Psilopogon malabaricus', 'Rhopocichla atriceps' = 'Dumetia atriceps', 'Chrysocolaptes lucidus' = 'Chrysocolaptes guttacristatus', 'Luscinia brunnea' = 'Larvivora brunnea', 'Parus aplonotus' = 'Machlolophus aplonotus', 'Turdoides striata' = 'Argya striata', 'Zoothera citrina' = 'Geokichla citrina', 'Turdoides subrufa' = 'Argya subrufa', 'Celeus brachyurus' = 'Micropternus brachyurus', 'Muscicapa ruficauda' = 'Ficedula ruficauda', 'Bubo nipalensis' = 'Ketupa nipalensis', 'Stigmatopelia chinensis' = 'Spilopelia chinensis','Acrocephalus aedon' = 'Arundinax aedon', 'Cyornis pallipes' = 'Cyornis pallidipes', 'Megalaima viridis' = 'Psilopogon viridis', 'Garrulax delesserti' = 'Pterorhinus delesserti', 'Acritillas indica' = 'Iole indica'))

vocal_act <- vocal_act %>% 
  left_join(sociality, by = 'scientific_name') 
```

## Peak frequency  

We prepare peak frequency as a predictor for the PGLS analysis
```{r}
# add standardized eBird codes to the frequency data
freq <-left_join (freq, species_codes[c(3,5)],
                  by = "species_annotation_codes")

# Only a total of 99 species are left after filtering species with very few templates
nTemplates_5 <- freq %>%
  group_by(eBird_codes) %>%
  count() %>%
  filter(n >= 5) %>%
  drop_na()

# left-join to remove species with less than 10 templates in the frequency dataset
freq_5 <- left_join(nTemplates_5[,1], freq)

# calculate median peak frequency after grouping by time of day and species
median_pf <- separate(freq_5, col = filename, into = c("site_id", "date", "time", "splits"), sep = "_") %>%
  mutate(
    time_of_day =
      case_when(time >= '060000' & time <= '100000' ~ 'dawn', 
                time >= '160000' & time <= '190000' ~ 'dusk')) %>%
  group_by(eBird_codes, time_of_day) %>%
  summarise(median_peak_freq =  median(peak_freq_in_Hz))

## join the frequency data to the vocal activity data
vocal_act <- vocal_act %>% 
  left_join(median_pf, by = c('eBird_codes', 'time_of_day')) %>%
  drop_na() 

# A total of 66 species were included and three species were excluded from final analysis.
```
 
## Trophic niche  

We prepare trophic niche categories for the PGLS analysis
```{r}
vocal_act <- vocal_act %>% 
  left_join(trait[,c(1,2,29)], by = c('scientific_name',
                          'common_name')) 

## remove species that are poorly represented by a particular trophic niche
vocal_act <- vocal_act %>%
  filter(trophic_niche != "Vertivore")

vocal_act <- vocal_act %>% mutate(trophic_niche = case_when (trophic_niche %in% 'Frugivore' ~ 'Frugivore, Granivore, Nectarivore', 
trophic_niche %in% 'Aquatic predator' ~ 'Frugivore, Granivore, Nectarivore', trophic_niche %in% 'Granivore' ~'Frugivore, Granivore, Nectarivore',
trophic_niche %in% 'Nectarivore' ~ 'Frugivore, Granivore, Nectarivore',
trophic_niche %in% 'Invertivore' ~'Invertivore',
trophic_niche %in% 'Omnivore' ~'Omnivore'))

vocal_act <- vocal_act %>% mutate(trophic_niche = case_when (trophic_niche %in% 'Invertivore' ~ '1', trophic_niche %in% 'Frugivore, Granivore, Nectarivore' ~ '2', trophic_niche %in% 'Omnivore' ~'3'))

## Only 62 species remain at this stage.
```

## Data cleaning

We clean the objects prior to running PGLS regressions
```{r}
# converting the scientific names of vocal_act as per tip.labels of the tree
vocal_act <- vocal_act %>% 
  mutate(across(scientific_name, str_replace, ' ', '_')) 
  
# converting the scientific names of vocal_act as per the birdtree data so that both are matching
vocal_act <- vocal_act %>%
mutate(scientific_name = 
         case_when (scientific_name == 'Dumetia_atriceps' ~ 'Rhopocichla_atriceps',                scientific_name == 'Chrysocolaptes_guttacristatus' ~ 'Chrysocolaptes_lucidus',
                 scientific_name == 'Turdus_simillimus' ~ 'Turdus_merula', 
                 scientific_name == 'Larvivora_brunnea' ~ 'Luscinia_brunnea',
                 scientific_name == 'Machlolophus_aplonotus' ~ 'Parus_xanthogenys', 
                scientific_name == 'Psilopogon_malabaricus' ~ 'Megalaima_rubricapillus',
                scientific_name == 'Tephrodornis_sylvicola' ~ 'Tephrodornis_gularis',
                scientific_name == 'Geokichla_citrina' ~ 'Zoothera_citrina',
               scientific_name == 'Cinnyris_asiaticus' ~ 'Nectarinia_asiatica',
               scientific_name == 'Leptocoma_minima' ~'Nectarinia_minima',
               scientific_name == 'Spilopelia_chinensis' ~'Stigmatopelia_chinensis',
               scientific_name == 'Argya_subrufa' ~ 'Turdoides_subrufa',
               scientific_name == 'Ficedula_nigrorufa' ~ 'Muscicapa_ruficauda',
               scientific_name == 'Gracula_indica' ~ 'Gracula_religiosa',
               scientific_name == 'Ketupa_nipalensis' ~ 'Bubo_nipalensis',
               scientific_name == 'Hypsipetes_ganeesa' ~ 'Hypsipetes_leucocephalus',
              scientific_name == 'Cyornis_pallidipes' ~ 'Cyornis_pallipes',
               scientific_name == 'Psilopogon_viridis' ~ 'Megalaima_viridis', .default = scientific_name)) #final dataframe with variables
         
# check same species in vocal_act and phylogenetic data, and pruning species from the tree which are not present in the vocal_act data
diff <- setdiff(tree$tip.label, vocal_act$scientific_name)
pruned.tree <- drop.tip (tree, diff) #final phylogenetic dataframe

setdiff(vocal_act$scientific_name,pruned.tree$tip.label) #checking if vocal_act and pruned.tree have the same species: result- 'character(0)'
```

## PGLS analysis
```{r}
## reference categories for categorical predictors/factors
## Territoriality: 1 = Non-territorial, 2 = Weakly territorial, 3 = Highly territorial
## Sociality: 0 = Non-communal signallers, 1 = Communal signallers
## Trophic niche: 1 = Invertivore, 2 = Frugivore, Granivore, Nectarivore, 3 = Omnivore

## Using normalized detections
model1 <- gls(percent_normalized_detections ~ scale(median_startTime) +
               relevel(factor(territory), ref = 3) + 
               factor(sociality) +
               scale(median_peak_freq) + 
               relevel(factor(trophic_niche), ref = 1), data = vocal_act, 
             correlation = corPagel(1, pruned.tree, form = ~scientific_name))

summary(model1)
plot_model(model1)

## Weakly territorial species showed a significant positive association with percent_normalized_detections 
## AIC: 1006.37

## Using just detections
model2 <- gls(percent_detections ~ scale(median_startTime) +
               relevel(factor(territory), ref = 3) + 
               factor(sociality) +
               scale(median_peak_freq) + 
               relevel(factor(trophic_niche), ref = 1), data = vocal_act, 
             correlation = corPagel(1, pruned.tree, form = ~scientific_name))

summary(model2)
plot_model(model2)

## Median start time showed a marginal positive association with percent detections
## AIC: 1004.313
```

########################################################################################

# Defining response variable- number of detections at each site and time
```{r}
# combining time_from_dawn and time_to_dusk as one column named 'time_from_startTime'
dawn <- acoustic_data %>%
  filter(time_of_day == "dawn") %>%
  dplyr::select(site_id, eBird_codes, detections, time_from_dawn, time_of_day, scientific_name) %>%
  rename(., time_from_startTime = time_from_dawn)
  
dusk <- acoustic_data %>%
  filter(time_of_day == "dusk") %>%
  dplyr::select(site_id, eBird_codes,detections, time_to_dusk, time_of_day, scientific_name) %>%
  rename(., time_from_startTime = time_to_dusk)

# combining the data and including vegetation data in this as well  
data <- bind_rows(dawn, dusk)
```

## Processing vegetation data
To use habitat type/vegetation data as a random effect in our model.
```{r}
## Process habitat structure variables
# Counting number of tree species and unique species per plot
treerich <- vegetation %>% 
  group_by(Site_ID) %>% 
  summarise (count = n(), richness = n_distinct(tree_species))

# Calculate average tree height across each unique site
treeheight <- vegetation %>% 
  drop_na(height) %>% 
  group_by(Site_ID) %>% 
  summarise(height = mean(height))

# Calculate basal area and left join with other data
basal_area <- vegetation %>% 
  mutate(basal_sum = rowSums(vegetation[,c(5:15)]^2)/(4*pi)) %>% group_by(Site_ID, Site_type) %>% 
  summarise(basal_area = sum(basal_sum)) 

# Calculate average canopy height
canopy_height <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(canopy_cover = mean(Canopy_cover))

# Calculate average leaf litter
leaf_litter <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(leaf_litter = mean(Leaf_litter))

# Calculate average vertical stratification
vert_strat <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(vert_strat = mean(Foliage_score))

# Year of planting
plantingYear <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(plantingYear = unique(Year.of.planting))

# Creating a final dataframe for further analysis
allVeg <- basal_area %>% 
  left_join(treeheight) %>%
  left_join(treerich) %>%
  left_join(canopy_height) %>%
  left_join(leaf_litter) %>%
  left_join(vert_strat) %>%
  left_join(plantingYear) %>%
  ungroup()

# converting the site_ids in vegetation_data as per the acoustic_data and renaming the column
allVeg <- allVeg %>% 
  mutate(across(Site_ID, str_remove_all,'_')) %>%
  rename(., site_id = Site_ID)

# joining the vegetation data
data <- data %>% 
  left_join(allVeg, by = 'site_id')
```

## Preparing the response and predictor variables
To use territoriality, sociality, peak freq, light availability, trophic niche as predictor variables.
```{r}
# filtering required columns
data <- data %>% 
              dplyr::select ('site_id', 'scientific_name', 'time_of_day', 'time_from_startTime',  'eBird_codes', 'detections', 'Site_type', 'basal_area', 'count', 'richness', 'canopy_cover', 'richness', 'leaf_litter', 'vert_strat') #here we have our response variable (detections), random effects (vegetation data, Site_type), and one of the predictor variables, i.e. light availability in the form of 'time_from_startTime'

# adding the rest of the predictor variables to the data
data <- data %>%
  left_join(territoriality, by = 'scientific_name') #including territoriality predictor variable with reference categories: 1= Non-territorial, 2= Weakly territorial, 3= Highly territorial

data <- data %>%
  left_join(sociality, by = 'scientific_name') #including sociality predictor variable with reference categories: 0= Non-communal signallers, 1= Communal signallers

data <- data %>% 
  left_join(trait[,c(1,2,29)], by = 'scientific_name') #including trophic niche predictor variable with reference categories: 1= Invertivore, 2= Frugivore, Granivore, Nectarivore, 3= Omnivore

data <- data %>% 
  left_join(median_pf, by = c('eBird_codes', 'time_of_day')) %>%
  drop_na() #including median peak frequency predictor variable
```

## Data cleaning

We clean the objects prior to running PGLMM
```{r}
# converting the scientific names of data as per tip.labels of the tree
data <- data %>% 
  mutate(across(scientific_name, str_replace, ' ', '_')) 
  
# converting the scientific names of data as per the birdtree data so that both are matching
data <- data %>%
mutate(scientific_name = 
         case_when (scientific_name == 'Dumetia_atriceps' ~ 'Rhopocichla_atriceps',                scientific_name == 'Chrysocolaptes_guttacristatus' ~ 'Chrysocolaptes_lucidus',
                 scientific_name == 'Turdus_simillimus' ~ 'Turdus_merula', 
                 scientific_name == 'Larvivora_brunnea' ~ 'Luscinia_brunnea',
                 scientific_name == 'Machlolophus_aplonotus' ~ 'Parus_xanthogenys', 
                scientific_name == 'Psilopogon_malabaricus' ~ 'Megalaima_rubricapillus',
                scientific_name == 'Tephrodornis_sylvicola' ~ 'Tephrodornis_gularis',
                scientific_name == 'Geokichla_citrina' ~ 'Zoothera_citrina',
               scientific_name == 'Cinnyris_asiaticus' ~ 'Nectarinia_asiatica',
               scientific_name == 'Leptocoma_minima' ~'Nectarinia_minima',
               scientific_name == 'Spilopelia_chinensis' ~'Stigmatopelia_chinensis',
               scientific_name == 'Argya_subrufa' ~ 'Turdoides_subrufa',
               scientific_name == 'Ficedula_nigrorufa' ~ 'Muscicapa_ruficauda',
               scientific_name == 'Gracula_indica' ~ 'Gracula_religiosa',
               scientific_name == 'Ketupa_nipalensis' ~ 'Bubo_nipalensis',
               scientific_name == 'Hypsipetes_ganeesa' ~ 'Hypsipetes_leucocephalus',
              scientific_name == 'Cyornis_pallidipes' ~ 'Cyornis_pallipes',
               scientific_name == 'Psilopogon_viridis' ~ 'Megalaima_viridis', .default = scientific_name)) #final dataframe with variables
         
# check same species in data and phylogenetic data, and pruning species from the tree which are not present in the data
diff <- setdiff(tree$tip.label, data$scientific_name)
pruned.tree <- drop.tip (tree, diff) #final phylogenetic dataframe

setdiff(data$scientific_name,pruned.tree$tip.label) #checking if data and pruned.tree have the same species: result- 'character(0)'
```

## PGLMM analysis
```{r}
# renaming site and sp columns
data <- data %>% 
  rename(., sp = scientific_name) %>%
  rename(., site = Site_type)
     
model <- pglmm(detections ~ scale(time_from_startTime) +  relevel(factor(territory), ref = 3) + factor(sociality) + scale(median_peak_freq) + relevel(factor(trophic_niche), ref = 1) + (1|site), data = data, family = "poisson",
                    cov_ranef = list(sp = pruned.tree))
summary (model)
 

```


