---
  editor_options: 
  chunk_output_type: console
---
  
  
# Sociality
  
In this script, we examine differences in vocal activity between dawn and dusk for each species in terms of their sociality. The expectation is that communal signallers would have a higher vocal activity, irrespective of dawn or dusk, than non-communal signallers.

Communal signallers are species that produce long-range acoustic signals in groups, either duets (two individuals) or choruses (three or more individuals) by both males and females. Non-communal signallers are species that sing as single individuals and not communally.

## Loading necessary libraries
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(vegan)
library(ggplot2)
library(scico)
library(data.table)
library(extrafont)
library(ggstatsplot)
library(ggpubr)
library(ggrepel)
library(ggdist)
library(gghalves)     
library(rstatix)
```

## Loading sociality data
```{r}
sociality <- read.csv("data/sociality-data.csv")

sociality <- sociality %>% dplyr::select (Species, Communal)
colnames (sociality) <- c("scientific_name", "sociality")

# Updating the scientific names of following species as per our dataset:
# Flame-throated bulbul- Rubigula gularis (earlier Pycnonotus gularis)
# Black eagle- Ictinaetus malaiensis (earlier Ictinaetus malayensis)
# Brown-capped pygmy woodpecker- Yungipicus nanus (earlier Dendrocopos nanus)
# Malabar barbet- Psilopogon malabaricus (earlier Megalaima malabarica)
# Dark-fronted babbler- Dumetia atriceps (earlier Rhopocichla atriceps)
# Greater flameback- Chrysocolaptes guttacristatus (earlier Chrysocolaptes lucidus)
# Indian blue robin- Larvivora brunnea (earlier Luscinia brunnea)
# Indian yellow tit- Machlolophus aplonotus (earlier Parus aplonotus)
# Jungle babbler- Argya striata (earlier Turdoides striata)
# Orange-headed thrush- Geokichla citrina (earlier Zoothera citrina)
# Rufous babbler- Argya subrufa (earlier Turdoides subrufa)
# Rufous woodpecker- Micropternus brachyurus (earlier Celeus brachyurus)
# Rusty-tailed flycatcher- Ficedula ruficauda (earlier Muscicapa ruficauda
# Spot-bellied eagle owl- Ketupa nipalensis (earlier Bubo nipalensis)
# Spotted dove- Spilopelia chinensis (earlier Stigmatopelia chinensis)
# Square-tailed bulbul- Hypsipetes ganeesa ganeesa (earlier Hypsipetes ganeesa)
# Thick-billed warbler- Arundinax aedon (earlier Acrocephalus aedon)
# White-bellied flycatcher- Cyornis pallidipes (earlier Cyornis pallipes)
# White-cheeked barbet- Psilopogon viridis (earlier Megalaima viridis)
# Wayanad laughingthrush- Pterorhinus delesserti (earlier Garrulax delesserti)
# Yellow-browed bulbul- Iole indica (earlier Acritillas indica)

sociality <- sociality %>% mutate(scientific_name = recode (scientific_name, 'Pycnonotus gularis' = 'Rubigula gularis', 'Ictinaetus malayensis' = 'Ictinaetus malaiensis', 'Dendrocopos nanus' = 'Yungipicus nanus', 'Megalaima malabarica' = 'Psilopogon malabaricus', 'Rhopocichla atriceps' = 'Dumetia atriceps', 'Chrysocolaptes lucidus' = 'Chrysocolaptes guttacristatus', 'Luscinia brunnea' = 'Larvivora brunnea', 'Parus aplonotus' = 'Machlolophus aplonotus', 'Turdoides striata' = 'Argya striata', 'Zoothera citrina' = 'Geokichla citrina', 'Turdoides subrufa' = 'Argya subrufa', 'Celeus brachyurus' = 'Micropternus brachyurus', 'Muscicapa ruficauda' = 'Ficedula ruficauda', 'Bubo nipalensis' = 'Ketupa nipalensis', 'Stigmatopelia chinensis' = 'Spilopelia chinensis', 'Hypsipetes ganeesa' = 'Hypsipetes ganeesa ganeesa', 'Acrocephalus aedon' = 'Arundinax aedon', 'Cyornis pallipes' = 'Cyornis pallidipes', 'Megalaima viridis' = 'Psilopogon viridis', 'Garrulax delesserti' = 'Pterorhinus delesserti', 'Acritillas indica' = 'Iole indica' ))
```

## Load acoustic data and species scientific names data
```{r}
acoustic_data <- read.csv("results/acoustic_data.csv")
species_codes <- read.csv("data/species-annotation-codes.csv")
```

## Vocal activity data
```{r}
# total number of acoustic detections summarized across every 10-s audio file
vocal_act <- acoustic_data %>%
  group_by(time_of_day, eBird_codes) %>%
  summarise(detections = sum(number)) %>%
  left_join(., species_codes[,c(1,2,5)], 
                       by = "eBird_codes") %>%
   group_by(eBird_codes) %>%
  mutate(total_detections =sum(detections)) %>%
  mutate(percent_detections = (detections/total_detections)*100) %>%
  ungroup()

# for further analysis, we remove species that were detected less than 20 times cumulatively across dawn and dusk
# this ensures that we keep species that called very few times in the morning and perhaps more times in the evening, or vice-versa

vocal_act <- vocal_act %>%
  filter(total_detections > 20)
```

## Join the vocal_activity data and sociality data
```{r}
vocal_act <- vocal_act %>% 
  left_join(sociality, by = 'scientific_name') %>%
  dplyr::mutate(sociality_cat = case_when (sociality %in% '0' ~ 'Non-communal signallers', sociality %in% '1' ~ 'Communal signallers'))
```

## Testing the differences using Wilcoxon test

Here, we see whether there are differences in the vocal activity between communal and non-communal signallers in dawn and dusk individually. 

```{r} 
stat.test <- vocal_act %>% 
  group_by (time_of_day) %>% 
  wilcox_test (percent_detections ~ sociality_cat)
```

There are 78 non-communal signallers and 21 communal signallers in dawn, and 69 non-communal and 20 communal signallers in dusk. 

## Visualization of % detections vs. sociality
```{r}
fig_soc_vocAct <- ggplot(vocal_act, aes(x = sociality_cat, y = percent_detections, fill = sociality_cat)) + 
   geom_boxplot(
    width = .2,
    size = 1.5, outlier.shape = NA
  )  +
  gghalves::geom_half_point(
    side = "l", 
    range_scale = .3, 
    alpha = .5, size = 3
  ) +
  labs(x = "", y = "Percent detections",
       fill = "Sociality Category") +
  facet_grid (~time_of_day) + 
  geom_signif(comparisons = list(c("Non-communal signallers", "Communal Signallers")), 
              map_signif_level = c("***"=0.001,"**"=0.01, "*"=0.05, " "=2),
              textsize = 5)+
  scale_color_manual(name = "Sociality", values = c("#0072B2", "#D55E00"),labels=c('Non-communal signallers', 'Communal Signallers'))+
  scale_fill_manual(name = "Sociality", values = c("#0072B2", "#D55E00"), labels=c('Non-communal signallers', 'Communal Signallers')) +
  theme_bw()+
  theme(legend.position = "none")+
  theme(text = element_text(family = "Century Gothic", size = 18, face = "bold"),plot.title = element_text(family = "Century Gothic",
      size = 18, face = "bold"),
      plot.subtitle = element_text(family = "Century Gothic", 
      size = 15, face = "italic",color="#1b2838"),
      axis.title = element_text(family = "Century Gothic",
      size = 18, face = "bold"))

ggsave(fig_soc_vocAct, filename = "figs/fig_sociality_vs_percentDetections.png", width = 14, height = 16, device = png(), units = "in", dpi = 300)
dev.off() 
```

There are no significant differences in % acoustic detections between communal and non-communal signllers in both dawn and dusk.
