
## Phylogenetic Linear Mixed Model (PGLMM) analysis
  
In this script, we examine which predictor/hypothesis (light availability, peak frequency, trophic niche, territoriality, and sociality) best explains the variation in vocal activity between dawn and dusk. We use linear regresssions while controlling for potential phylogenetic signals/relationships, and adding site_type as random effect.

[Note: if you want to run this script, please run sript corresponding to pgls-analysis to load all objects required]

# Defining response variable- number of detections at each site and time
```{r}
# combining time_from_dawn and time_to_dusk as one column named 'time_from_startTime'
dawn <- acoustic_data %>%
  filter(time_of_day == "dawn") %>%
  dplyr::select(site_id, eBird_codes, detections, time_from_dawn, time_of_day, scientific_name) %>%
  rename(., time_from_startTime = time_from_dawn)
  
dusk <- acoustic_data %>%
  filter(time_of_day == "dusk") %>%
  dplyr::select(site_id, eBird_codes,detections, time_to_dusk, time_of_day, scientific_name) %>%
  rename(., time_from_startTime = time_to_dusk)

# combining the data and including vegetation data in this as well  
data <- bind_rows(dawn, dusk)
```

## Processing vegetation data
To use habitat type/vegetation data as a random effect in our model.
```{r}
## Process habitat structure variables
# Counting number of tree species and unique species per plot
treerich <- vegetation %>% 
  group_by(Site_ID) %>% 
  summarise (count = n(), richness = n_distinct(tree_species))

# Calculate average tree height across each unique site
treeheight <- vegetation %>% 
  drop_na(height) %>% 
  group_by(Site_ID) %>% 
  summarise(height = mean(height))

# Calculate basal area and left join with other data
basal_area <- vegetation %>% 
  mutate(basal_sum = rowSums(vegetation[,c(5:15)]^2)/(4*pi)) %>% group_by(Site_ID, Site_type) %>% 
  summarise(basal_area = sum(basal_sum)) 

# Calculate average canopy height
canopy_height <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(canopy_cover = mean(Canopy_cover))

# Calculate average leaf litter
leaf_litter <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(leaf_litter = mean(Leaf_litter))

# Calculate average vertical stratification
vert_strat <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(vert_strat = mean(Foliage_score))

# Year of planting
plantingYear <- vegetation %>%
  group_by(Site_ID) %>%
  summarise(plantingYear = unique(Year.of.planting))

# Creating a final dataframe for further analysis
allVeg <- basal_area %>% 
  left_join(treeheight) %>%
  left_join(treerich) %>%
  left_join(canopy_height) %>%
  left_join(leaf_litter) %>%
  left_join(vert_strat) %>%
  left_join(plantingYear) %>%
  ungroup()

# converting the site_ids in vegetation_data as per the acoustic_data and renaming the column
allVeg <- allVeg %>% 
  mutate(across(Site_ID, str_remove_all,'_')) %>%
  rename(., site_id = Site_ID)

# joining the vegetation data
data <- data %>% 
  left_join(allVeg, by = 'site_id')
```

## Preparing the response and predictor variables
To use territoriality, sociality, peak freq, light availability, trophic niche as predictor variables.
```{r}
# filtering required columns
data <- data %>% 
              dplyr::select ('site_id', 'scientific_name', 'time_of_day', 'time_from_startTime',  'eBird_codes', 'detections', 'Site_type', 'basal_area', 'count', 'richness', 'canopy_cover', 'richness', 'leaf_litter', 'vert_strat') #here we have our response variable (detections), random effects (vegetation data, Site_type), and one of the predictor variables, i.e. light availability in the form of 'time_from_startTime'

# adding the rest of the predictor variables to the data
data <- data %>%
  left_join(territoriality, by = 'scientific_name') #including territoriality predictor variable with reference categories: 1= Non-territorial, 2= Weakly territorial, 3= Highly territorial

data <- data %>%
  left_join(sociality, by = 'scientific_name') #including sociality predictor variable with reference categories: 0= Non-communal signallers, 1= Communal signallers

data <- data %>% 
  left_join(trait[,c(1,2,29)], by = 'scientific_name') #including trophic niche predictor variable with reference categories: 1= Invertivore, 2= Frugivore, Granivore, Nectarivore, 3= Omnivore

data <- data %>% 
  left_join(median_pf, by = c('eBird_codes', 'time_of_day')) %>%
  drop_na() #including median peak frequency predictor variable
```

## Data cleaning

We clean the objects prior to running PGLMM
```{r}
# converting the scientific names of data as per tip.labels of the tree
data <- data %>% 
  mutate(across(scientific_name, str_replace, ' ', '_')) 
  
# converting the scientific names of data as per the birdtree data so that both are matching
data <- data %>%
mutate(scientific_name = 
         case_when (scientific_name == 'Dumetia_atriceps' ~ 'Rhopocichla_atriceps',                scientific_name == 'Chrysocolaptes_guttacristatus' ~ 'Chrysocolaptes_lucidus',
                 scientific_name == 'Turdus_simillimus' ~ 'Turdus_merula', 
                 scientific_name == 'Larvivora_brunnea' ~ 'Luscinia_brunnea',
                 scientific_name == 'Machlolophus_aplonotus' ~ 'Parus_xanthogenys', 
                scientific_name == 'Psilopogon_malabaricus' ~ 'Megalaima_rubricapillus',
                scientific_name == 'Tephrodornis_sylvicola' ~ 'Tephrodornis_gularis',
                scientific_name == 'Geokichla_citrina' ~ 'Zoothera_citrina',
               scientific_name == 'Cinnyris_asiaticus' ~ 'Nectarinia_asiatica',
               scientific_name == 'Leptocoma_minima' ~'Nectarinia_minima',
               scientific_name == 'Spilopelia_chinensis' ~'Stigmatopelia_chinensis',
               scientific_name == 'Argya_subrufa' ~ 'Turdoides_subrufa',
               scientific_name == 'Ficedula_nigrorufa' ~ 'Muscicapa_ruficauda',
               scientific_name == 'Gracula_indica' ~ 'Gracula_religiosa',
               scientific_name == 'Ketupa_nipalensis' ~ 'Bubo_nipalensis',
               scientific_name == 'Hypsipetes_ganeesa' ~ 'Hypsipetes_leucocephalus',
              scientific_name == 'Cyornis_pallidipes' ~ 'Cyornis_pallipes',
               scientific_name == 'Psilopogon_viridis' ~ 'Megalaima_viridis', .default = scientific_name)) #final dataframe with variables
         
# check same species in data and phylogenetic data, and pruning species from the tree which are not present in the data
diff <- setdiff(tree$tip.label, data$scientific_name)
pruned.tree <- drop.tip (tree, diff) #final phylogenetic dataframe

setdiff(data$scientific_name,pruned.tree$tip.label) #checking if data and pruned.tree have the same species: result- 'character(0)'
```

## PGLMM analysis
```{r}
# renaming site and sp columns
data <- data %>% 
  rename(., sp = scientific_name) %>%
  rename(., site = Site_type)
     
model <- pglmm(detections ~ scale(time_from_startTime) +  relevel(factor(territory), ref = 3) + factor(sociality) + scale(median_peak_freq) + relevel(factor(trophic_niche), ref = 1) + (1|site), data = data, family = "poisson",
                    cov_ranef = list(sp = pruned.tree))
summary (model)
```


